### 2018-06-01 source/Ch2-intro/calibration.tex
added a lot

\pg
The distinction between these two noise sources is crucial; one can never go `below noise' for a given observation, no matter the quality of calibration. Astronomers typically observe for longer periods of time in order to reduce $\sigma_{thermal}$ in their images, but this will not reduce the artefacts caused by poor calibration solutions. Uncorrected Direction-Dependent Effects will not go away on their own, no matter how long the integration time. Similarly, there is a limit to how much improving calibration will improve the final image: eventually, more data is required to drive noise down.

@@

\pg
While values for $a_p$ and $b_p$ can in theory be found for both autocorrelation and both crosscorrelations, low signal-to-noise means that in practice, a single set of values is solved for per antenna\footnote{This reduces calibration to solving only for the intensity gains: the data can then only be used for \emph{intensity mapping} (e.g. \citepads{1957IAUS....4..159J}). This practice therefore precludes polarimetry.}. With these techniques, one can achieve dynamic ranges of about 100:1 (\citepads{2010A&A...524A..61N}).

@@

\pg
By extending this idea to VLBI\footnote{e.g.  \citepads{1974ApJ...193..293R}}, \emph{amplitude closure} was introduced to the field along with phase closure \citepads[as described in][]{1983Sci...219...51R}. These quantities are immune to antenna-based effects.%Astronomers were quick to apply these methods to interferometers in general.

@@

\pg
In this section, I will discuss the approach taken by \textcolor{red}{[cite DDF and kMS paper when it comes out]} and \textcolor{red}{[cite prefactor paper if/when it is out]}, \emph{facet-based calibration}. This method has an imaging equivalent \citepads[see][and associated papers]{2017arXiv171202078T}. The approach cannot be divorced from imaging, for the simple reason that solving for and applying Jones matrices in different directions requires image-plane knowledge of the sky brightness distribution $\Bmatrix$. % The imaging challenges this methods introduces will be discussed in [href to relevant imaging section].

@@

\pg
Why then not solve for a set of calibration solutions for each source in the calibration model, iteratively adding new sources to this model as they become above the noise as calibration artefacts - introduced by poor calibration - decrease? There are two main reasons for this: firstly, the signal-to-noise when calibrating using very faint source models will be poor. This means that gains will likely be extremely noisy, and applying them will result in increased calibration artefacts in the final image. Secondly, even by optimising for signal-to-noise and only using a few of the brightest sources in the field, the problem quickly becomes intractably large in terms of computing power required, provided that standard complex differentiation is used \citepads[see][]{2016ApJS..223....2V}. Wirtinger differentiation \citepads[see][and associated papers]{2014arXiv1410.8706T} does alleviate this somewhat, but is not implemented in all standard direction-dependent calibration software at the time of writing.

>>>>

\pg
The distinction between these two noise sources is crucial; one can never go `below noise' for a given observation, no matter the quality of calibration. Astronomers typically observe for longer periods of time in order to reduce $\sigma_{thermal}$ in their images, but this will not reduce the artefacts caused by poor calibration solutions. Uncorrected Direction-Dependent Effects\footnote{See \cref{section.RIME.MultiplePointSources}} will not go away on their own, no matter how long the integration time. Similarly, there is a limit to how much improving calibration will improve the final image: eventually, more data is required to drive noise down.

@@

\pg
While values for $a_p$ and $b_p$ can in theory be found for both autocorrelation and both crosscorrelations (e.g. XY and YX for linear receptors), low signal-to-noise means that in practice, a single set of values is solved for per antenna\footnote{This reduces calibration to solving only for the intensity gains: the data can then only be used for \emph{intensity mapping} (e.g. \citepads{1957IAUS....4..159J}). This practice therefore precludes polarimetry.}. With these techniques, one can achieve dynamic ranges of about 100:1 (\citepads{2010A&A...524A..61N}).

@@

\pg
By extending this idea to VLBI (see e.g.  \citepads{1974ApJ...193..293R}), \emph{amplitude closure} was introduced to the field along with phase closure \citepads[as described in][]{1983Sci...219...51R}. These quantities are immune to antenna-based effects.%Astronomers were quick to apply these methods to interferometers in general.

@@

\pg
In this section, I will discuss the \emph{facet-based} approach taken by \citetads{2017arXiv171202078T}. The approach cannot be divorced from imaging, for the simple reason that solving for and applying Jones matrices\footnote{See \cref{chapter.RIME} for a full discussion on calibration and Jones matrices} in different directions requires image-plane knowledge of the sky brightness distribution $\Bmatrix$. % The imaging challenges this methods introduces will be discussed in [href to relevant imaging section].

@@

\pg
Why then not solve for a set of calibration solutions for each source in the calibration model, iteratively adding new sources to this model as they become above the noise as calibration artefacts - introduced by poor calibration - decrease? There are two main reasons for this: firstly, the SNR when calibrating using very faint source models will be poor. This means that gains will likely be extremely noisy, and applying them will result in increased calibration artefacts in the final image. Secondly, even by optimising for signal-to-noise and only using a few of the brightest sources in the field, the problem quickly becomes intractably large in terms of computing power required, provided that standard complex differentiation is used \citepads[see][]{2016ApJS..223....2V}. Wirtinger differentiation \citepads[see][]{2014arXiv1410.8706T} does alleviate this somewhat by allowing the Hessian matrix to be written as a block-diagonal matrix, but is not implemented in all standard direction-dependent calibration software at the time of writing.

### 2018-06-01 source/Ch2-intro/imaging.tex
added a lot

\pg
However, because radio frequencies are so much lower, achieving a resolution comparable to e.g. the HST requires extremely large dishes. While there exist telescopes, both old and new, which work on this principle (from Arecibo Observatory, shown in \cref{fig.arecibo}, to the upcoming FAST telescope in China, shown in \cref{fig.FAST})

@@

\pg
Calibrating radio antennas is usually a relatively straightforward matter - the signal loss and distortion as the dish converts the electromagnetic wave to a voltage can be described, based on the antenna, either as a simple scaling factor, or a complex number (giving information on phase and amplitude errors). This model of antenna loss and distortion is referred to as the \emph{antenna gain}. Solving for the antenna gains used in more complex interferometric array is a problem described in Section \ref{section.calibration}.

\pg
For now, we will assume that calibration has been performed perfectly, and that the gain-corrections have been applied to the voltage measured by our antennas. We therefore work with the true signal from astrophysical sources.

@@

\pg
There are two quantities of interest to all astronomers: sensitivity and resolution. An instrument's sensitivity is a function of its collecting area\footnote{It is also a function of technological factors, of course, but \emph{ceteris paribus}, a more sensitive telescope means a telescope with a wider collecting area}. Resolution, for well-designed instruments\footnote{By this, we mean that we assume that an instrument is also designed to optimise resolution.}, is limited by diffraction. This introduces specific issues in the radio domain. Radio waves have very long wavelengths - often comparable to meters, rather than the 100$nm$ wavelengths of optical light. Achieving a resolution comparable to those of optical telescopes would thus require making telescopes with apertures tens of millions of times larger than those already titanic instruments!

\pg
In practice, this technical constraint on resolution is one that astronomers would be very interested in overcoming. This technical limitation therefore demands a technical solution. In practice, this solution consists of recourse to interferometric techniques. Indeed, interferometry can be thought of as the construction of a "sparse" dish, as illustrated in Fig.

\begin{figure}[ht]
\centering
\begin{subfigure}{.43\textwidth}
\resizebox{\hsize}{!}{\includegraphics{images/baseline-resolution.png}}
\caption{\label{fig.baseline.image} A pair of dishes can surpass the resolution limit of its components.}
\end{subfigure}
\hfill
\begin{subfigure}{.43\textwidth}
\resizebox{\hsize}{!}{\includegraphics{images/arecibo-vla.png}}
\caption{\label{fig.arecibo.vla} With enough pairs of dishes, it is possible to synthesise a much larger dish.}
\end{subfigure}
\caption{\label{fig.aperture.synthesis} Illustration of the underlying principle of interferometry. The 27 dishes of the VLA can be thought of as "synthesising" a similar circular dish as Arecibo. This idea is the reason why radio interferometry is historically known as "aperture synthesis" in the literature of radio astronomy. Both images are copyrighted by Addison Wesley. \textcolor{red}{CHANGE FIG B. AS THE ELLIPSE IS NOT SEE-THROUGH}}
\end{figure}

\pg
Of course, this resolution improvement does not come for free. To understand the cost of interferometry, we must discuss the properties of its core component: the baseline.

@@

\subsection{The Visibility}

\pg
We have defined what a baseline corresponds to: a vector coordinate in $(u,v,w)$-space. To each baseline we associate a measurement, which we call the \emph{visibility}. 
\begin{figure}[ht]
\centering
\includegraphics[width=.7\textwidth]{images/visibility-creation.png}
\caption{\label{fig.visibility} There are two ways to combine the voltages from two antennas into a visibility: they are sum-correlation and $\pi$-correlation. In this manuscript, we will only concern ourselves with the latter. Image credit: Julien Girard}
\end{figure}
The visibility associated with baseline $\mathbf{b}_{AB}$ is created by taking the voltage measured by antenna A, multiply it by the complex conjugate of the voltage measured by antenna B, average over the correlator dump time (i.e. the time over which the measurement is made). This scalar quantity is then multiplied by the baseline position vector. In other words:
\begin{equation}
\mathbf{b}_{AB} = \langle V_{A} V_{B}^*\rangle_{\delta t, \delta \nu} \frac{\mathbf{x}_{B}-\mathbf{x}_{A}}{\nu_\mathrm{obs}}
\end{equation}
where $\langle \cdots \rangle_{x}$ denotes an averaging over length $x$. We see that a visibility is a complex vector quantity. We also see that $\mathbf{b}_{AB} = \mathbf{b}_{BA}^*$: the information of the visibility associated with baseline BA is contained in the visibility associated with baseline AB. This means that in practice, only half of the visibilities ever need be stored. What does the visibility measurement correspond to?
\begin{figure}[ht]
\centering
\includegraphics[width=.5\textwidth]{images/visibility-measure.png}
\caption{\label{fig.visibility.measure} Here, we assume that there are only two sources in the sky whose signal can be measured by our antennas. The final visibility is the sum of the visibilities associated with each individual source. Image credit: Julien Girard}
\end{figure}

\pg
Different astrophysical sources are located very far apart, and so their signals do not interfere. They are measured additively in both antennas. Provided that the signal from both sources is coherent when observed by the dishes, the correlation between the voltages measured by antennas A and B will simply be the sum of the voltage correlations associated with individual sources. % - i.e. the interferometric signal from different sources are additive.

\pg
Note that in Fig. \ref{fig.visibility.measure}, neither source is at the zenith. This introduces the notion of the \emph{effective baseline} which is shorter than the \emph{physical baseline} when measuring a signal from a source that is not at the zenith. Sources in different positions in the sky will be measured by an interferometric array, but the same array will have different effective baselines for different sources. To point an interferometric array in a specific direction in the sky, one must digitally introduce a phase delay in each antenna (or, in older interferometers, by playing with the cable length between each dish and the correlator). This ensures that the constituent antennas' sensitivities are maximised in the direction of interest.

@@

\pg
In general, interferometric design is such that the $w$ component of visibilities' $(u,v,w)$ coordinates is negligible (or can be put in a frame of reference where it can usually be approximated as such). Radio astronomers tend to thus talk of a $uv$-plane rather than $uvw$-space to describe visibility space. The set of $uv$-values for all the baselines of an interferometric array is known as its $uv$-coverage, and defines the array's properties entirely.

\pg
For the VLA, for example, the $uv$-coverage when observing the zenith will be as shown in Fig. \ref{fig.vla.uvcoverage}.

@@

\pg
Assuming that all the antennas in an array are equivalent and perfectly calibrated, the van Cittert-Zernike theorem (\citetads{1934Phy.....1..201V}, covered in \citetads{2001isra.book.....T}) allows us to equate a visibility with a single Fourier mode of the plane tangent to the sky where the instrument is pointed. This direction is called the \emph{phase centre}, because digital phase shifts are introduced between antenna voltages before averaging so as to point each visibility in this direction.

\pg
The Point-Spread Function (PSF) of an array is thus the Fourier transform of the array's PSF. The more element in the array, the better the PSF. However, an interferometer's PSF will always be worse than an equivalently-sized antenna dish, the UV-coverage of which is the inverse Fourier transform of a disc. This is because an interferometer behaves like a \emph{sparse} dish: the sparsity of our coverage manifests as very large point-spread function. The more baselines are available, the less sparse the UV-coverage, and so the better the PSF. This is shown in Fig. \ref{fig.vla.PSFvsUVcov}.
\begin{figure}[h!]
\centering
\resizebox{\hsize}{!}{\includegraphics{images/PSFvsUVcov.png}}
\caption{\label{fig.vla.PSFvsUVcov}Plot of the PSF associated to more and more complete UV-coverage (here, consisting of more and more elements of the VLA). Image credit: Rick Perley, 3GC4 lecture.}
\end{figure}

\pg
This PSF can be improved by assuming that the sky is constant over periods of time (supersynthesis) or constant over fractions of the total bandwidth. This improves UV-coverage by a factor $N_\mathrm{t}\times N_\nu$. The effect of supersynthesis on the PSF is shown in Fig. \ref{fig.vla.PSFsupersynthesis}.
\begin{figure}[h!]
\centering
\resizebox{\hsize}{!}{\includegraphics{images/{PSF_supersynthesis}.png}}
\caption{\label{fig.vla.PSFsupersynthesis}Effect of supersynthesis on the UV-coverage of the full VLA. Image credit: Rick Perley, 3GC4 lecture.}
\end{figure}

@@

\pg
In practice, performing a full deconvolution on large images (easily over 25 million pixels in total) is not computationally viable. Scientists therefore resort to algorithms and techniques to accelerate imaging and deconvolution. For example, the use of Direct Fourier Transforms (DFT), which are extremely slow, is avoided when going from visibility space to image space. Similarly, one seeks to avoid having to perform full deconvolution on the dirty image, since many pixels in the image do not actually contain astrophysical signal. In this chapter, we will briefly cover the most common method of deconvolution used in radio astronomy.

@@

\pg
In Fig. \ref{fig.vla.clean}, we show the effect of deconvolution on a dirty image of 3C295 (a very bright \& complex source which is the subject of much of this thesis' work). Since this source includes diffuse emission, standard CLEAN is insufficient (since CLEAN assumes a sparse distribution of point-like sources). Instead, DDFacet's novel MetroCLEAN algorithm (see \citetads{2017arXiv171202078T} for more details), which deconvolves islands consisting of sets of pixels at once, was used. This allowed for a far greater dynamic range than standard CLEAN could achieve.

\pg
The problem of imaging radio interferometric data, of which deconvolution is but a subset, is still a very active field of research today. For more information on this topic, the NRAO website\footnote{See \href{https://www.cv.nrao.edu/~abridle/deconvol/node8.html\#SECTION00051000000000000000}{the CLEAN section of the NRAO website}.} is an excellent start for a general introduction to this topic, while \citet{2017arXiv171202078T} provides an outstanding entry point for cutting-edge work at the time of writing..

\section{Weighting Schemes}
\pg
In \cref{section.clean}, we discuss the problem of deconvolution in radio interferometric imaging. One consideration that is not raised is the issue of conditioning. Consider two extreme cases: a field containing a single point source, and a field containing a diffuse and turbulent source, with very complex flux distribution at all scales. These cases are shown in \cref{fig.point,fig.cyga} respectively. Deconvolving the PSF from the first image using CLEAN algorithms will be child's play, but doing the same with the second will be extremely complex. The first case is said to be well-conditioned, and the second to be poorly-conditioned.

@@

\pg
The better the conditioning, the easier the deconvoution and the closer to the ground truth its result. As such, when the true underlying structure of a source is poorly-known, getting the best conditioning possible for deconvolution becomes a key concern. Thankfully, because the actual measurements taken with interferometers are the Fourier modes of the sky brightness distribution, they can be weighted when reconstructing the images to fulfill specific requirements. In particular, Briggs weighting \citepads{1995AAS...18711202B} gives a sliding parameter between \emph{natural} weighting, where each visibility is weighed equivalently (this optimises noise in the field) and \emph{uniform} weighting, which gives each visibility a weight inversely proportional to $uv$-plane density. Uniform weighting optimises for resolution at the cost of image signal-to-noise.

\begin{figure}[h!]
\centering
\begin{subfigure}{.48\textwidth}\resizebox{\hsize}{!}{\includegraphics{images/natural.png}}
\caption{\label{fig.natural} Image made with LOFAR data and natural weighting.}
\end{subfigure}
\hfill
\begin{subfigure}{.48\textwidth}
\resizebox{\hsize}{!}{\includegraphics{images/uniform.png}}
\caption{\label{fig.uniform} Image made with the same data as \cref{fig.natural} and uniform weighting.}
\end{subfigure}
\hfill
\begin{subfigure}{.48\textwidth}
\resizebox{\hsize}{!}{\includegraphics{images/briggs0.png}}
\caption{\label{fig.briggs0} Image made with the same data as \cref{fig.natural} and Briggs weighting.}
\end{subfigure}
\caption{\label{fig.briggs-weighting}Impact of different weighting schemes on interferometric image reconstruction. All the images shown above are deconvolved.}
\end{figure}

@@

\pg
Note that the weighting schemes shown here rely only on a choice between normalising visibilities by local $uv$-density vs. signal-to-noise, with Briggs weighting providing a formal method for going from one regime to another or optimising between both. Other weighting schemes exist - one can attempt to optimise data size vs decorrelation\footnote{As data is averaged, decorrelation becomes more and more important in the field. See \cref{section.RIME} for more details.}, for example \citepads[see][and related work]{2016MNRAS.462.2542A}. This can help ensure a homogeneous conditioning of the deconvolution problem in the image, by ensuring that the PSF is as peaked as possible throughout the field of interest.

\pg
Finally, one type of weighting scheme of note is simple data flagging \citepads[cf.]{2010ascl.soft10017O}. This consists of assigning a null weight to those data points considered too corrupted by various instrumental or atmospheric effects (typically, Radio Frequency Interference - RFI) to be scientifically useful. Those weights which are not deemed too corrupted are assigned a unit weight. This corresponds to simply dropping those data points which would corrupt the image reconstruction by introducing unphysically strong fringes in the image.

>>>>

\pg
However, because radio frequencies are so much lower, achieving a resolution comparable to e.g. the HST requires extremely large dishes. While there exist telescopes, both old and new, which work on this principle (from Arecibo Observatory, shown in \cref{fig.arecibo}, to the upcoming FAST telescope in China, shown in \cref{fig.FAST}), the associated technical difficulties (pointing is complex business, as are the associated optics requirements; maintenance costs are high, etc...) have made the single-dish approach prohibitive.

@@

\section{An Observer's Perspective of Radio Interferometry}

\pg
Astronomers who request observation time on interferometric arrays are generally not experts in the theory and operation of these arrays. This is a simple consequence of scientific division of labour. This thesis is written from the perspective of an ``expert" user: one who specialises in understanding interferometric arrays and reducing their data, rather than their subsequent analysis. However, it can be very helpful to contextualise this perspective in the experience of other scientists, which is what this section aims to do.

\pg
LOFAR data can be acquired through observation or via the Long-Term Archive. The worst of this data will already be flagged. Users can thus generally proceed directly to calibration. Because there is no guarantee of sufficient signal-to-noise in the ``target field" (the object that the astronomer is actually interested in), it is standard practice to observe a calibrator source for 5 minutes before and after each observation. Such sources typically need to be bright and unresolved for the interferometer being used. Because they are bright and at phase centre for the 5-minute observation, the SNR when solving for calibration solutions for these 5 minutes will tend to be quite good. These solutions will then be interpolated between the 5 minutes onto the 8-hour observation. Calibration will be described in \cref{section.calibration}.

\pg
So far, the astronomer will have worked entirely with visibilities. However, generally speaking, the aim will be to get information on the object's brightness distribution at the observing frequency. This will require \emph{imaging}: going from visibilities to the image-plane. This can be done through a variety of tools, but will generally involve deconvolution. Imaging will be the subject of the rest of this chapter, and deconvolution will be described in more detail in \cref{section.clean}.

\pg
After obtaining the initial image, it may be desirable to perform a few rounds of self-calibration \citepads[cf.]{2018arXiv180505266B} on the target field. This consists of extracting a model from the initial image and using it as a new calibration model, then re-imaging. Doing this will usually dramatically improve the final image.

\pg
The framework for interferometric calibration is complex, and understanding it requires a good grasp on interferometry itself. For now, we will assume that calibration has been performed perfectly, and that the gain-corrections have been applied to the voltage measured by our antennas. We therefore work with the true signal from astrophysical sources until \cref{section.RIME}.

@@

\pg
There are two quantities of interest to all astronomers: sensitivity and angular resolution\footnote{Other kinds of resolutions - in time and frequency, for example - are also extremely important to many astronomers, but are not what concern us here.}. An instrument's sensitivity is a function of its collecting area\footnote{It is also a function of technological factors, of course, but \emph{ceteris paribus}, a more sensitive telescope means a telescope with a wider collecting area}. Resolution, for well-designed instruments\footnote{By this, we mean that we assume that an instrument is also designed to optimise resolution.}, is limited by diffraction. This introduces specific issues in the radio domain. Radio waves have very long wavelengths - often comparable to meters, rather than the $\sim100-1000$nm wavelengths of optical light. Achieving a resolution comparable to those of optical telescopes would thus require making telescopes with apertures tens of millions of times larger than those already titanic instruments!

\pg
In practice, this technical constraint on resolution is one that astronomers are very interested in overcoming. This technical limitation therefore demands a technical solution. In practice, this solution consists of recourse to interferometric techniques. Indeed, interferometry can be thought of as the construction of a "sparse" dish, as illustrated in \cref{fig.aperture.synthesis}.

\begin{figure}[ht]
\centering
\begin{subfigure}{.43\textwidth}
\resizebox{\hsize}{!}{\includegraphics{images/baseline-resolution.png}}
\caption{\label{fig.baseline.image} A pair of dishes can surpass the resolution limit of its components.}
\end{subfigure}
\hfill
\begin{subfigure}{.43\textwidth}
\resizebox{\hsize}{!}{\includegraphics{images/sparseArray.png}}
\caption{\label{fig.arecibo.vla} With enough pairs of dishes, it is possible to synthesise a much larger dish.}
\end{subfigure}
\caption{\label{fig.aperture.synthesis} Illustration of the underlying principle of interferometry. The 27 dishes of the VLA can be thought of as "synthesising" a similar circular dish as Arecibo. This idea is the reason why radio interferometry is historically known as "aperture synthesis" in the literature of radio astronomy. \cref{fig.baseline.image} is copyrighted by Addison Wesley.} % \textcolor{red}{CHANGE FIG B. AS THE ELLIPSE IS NOT SEE-THROUGH}}
\end{figure}

\pg
Interferometric observations follow a very familiar pattern: the elements of the interferometric array (which can be single dishes or phased arrays\footnote{Described in \cref{sec.visibility}} of smaller antennas) measure a voltage. This voltage is correlated between antennas, as will be described shortly. Each of these correlations correspond to a single Fourier mode of the sky, as per the Zernike-van Cittert theorem (cf. \cref{sec.imag.psf}). They must be calibrated, by comparing the observations of a calibrator source and modelled visibilities. Calibration is described in much greater detail in \cref{sec.RIME}. Then, the calibrated visibilities are used to reconstruct an image of the sky, which often requires deconvolution due to the sparsity of interferometric arrays.

\pg
The resolution improvement of interferometers does thus not come for free. To better understand the cost of interferometry, we will cover calibration after explaining the particularities of interferometers. We will begin with the properties of their core component: the baseline.

@@

\subsection{The Visibility}\label{sec.visibility}

\pg
We have defined what a baseline corresponds to: a vector coordinate in $(u,v,w)$-space. To each baseline we associate a measurement, which we call the \emph{visibility}. 
\begin{figure}[ht]
\centering
\includegraphics[width=.7\textwidth]{images/visibility-creation.png}
\caption{\label{fig.visibility} There are two ways to combine the voltages from two antennas into a visibility: they are sum-correlation and $\pi$-correlation. In this manuscript, we will only concern ourselves with the latter. Image credit: Julien Girard}
\end{figure}
The visibility associated with baseline $\mathbf{b}_{AB}$ is created by taking the voltage measured by antenna A, multiply it by the complex conjugate of the voltage measured by antenna B, average over the correlator dump time (i.e. the time over which the measurement is made). This scalar quantity is associated to the baseline position vector $\mathbf{b}_{AB}$. In other words:
\begin{align}
\mathbf{b}_{AB} &= \frac{\mathbf{x}_{B}-\mathbf{x}_{A}}{\lambda_\mathrm{obs}}\\
V_{AB}          &= \langle V_{A} V_{B}^*\rangle_{\delta t, \delta \nu} 
\end{align}
where $\langle \cdots \rangle_{i}$ denotes an averaging over quantity $i$. We see that a visibility is a complex vector quantity. We also see that $\mathbf{b}_{AB} = \mathbf{b}_{BA}^*$: the information of the visibility associated with baseline BA is contained in the visibility associated with baseline AB. This means that in practice, only half of the visibilities ever need be stored. What does the visibility measurement correspond to?
\begin{figure}[ht]
\centering
\includegraphics[width=.5\textwidth]{images/visibility-measure.png}
\caption{\label{fig.visibility.measure} Here, we assume that there are only two sources in the sky whose signal can be measured by our antennas. The final visibility is the sum of the visibilities associated with each individual source. Image credit: Julien Girard \textcolor{red}{add reference to julien's phd}}
\end{figure}

\pg
Different astrophysical sources are located very far apart, and so their signals do not interfere. They are measured additively in both antennas. Provided that the signal from both sources is coherent when observed by the dishes, the correlation between the voltages measured by antennas A and B will simply be the sum of the voltage correlations associated with individual sources. This is shown in \cref{fig.visibility}: the radio signal from two different sources, at two different locations, is picked up simultaneously by both radio antennas. Because only the correlation between both antennas is measured, the signals do not interfere. % - i.e. the interferometric signal from different sources are additive.
\pg
Note that in Fig. \ref{fig.visibility.measure}, neither source is at the zenith. This introduces the notion of the \emph{effective baseline} which is shorter than the \emph{physical baseline} when measuring a signal from a source that is not at the zenith. Sources in different positions in the sky will be measured by an interferometric array, but the same array will have different effective baselines for different sources. To point an interferometric array in a specific direction in the sky, one must digitally introduce a phase delay in each antenna (or, for fundamental interferometer elements such as the LOFAR tiles or NenuFAR, by playing with the cable length between each dish and the correlator). This ensures that the constituent antennas' sensitivities are maximised in the direction of interest. The position of sources are then given in terms of the cardinal angle $\sigma$: this is the difference between the position of a source and the phase centre, which is where the array is pointing. For example, in \cref{fig.visibility.measure}, if we point antennas and introduce phase delays such that we are pointing the array towards Source 2, the position of Source 1 will be $\sigma$.

@@

\pg
In general, interferometric design is such that the $w$ component of visibilities' $(u,v,w)$ coordinates is negligible (or can be put in a frame of reference where it can usually be approximated as such). Radio astronomers tend to thus talk of a $uv$-plane rather than $uvw$-space to describe visibility space. The set of $uv$-values for all the baselines of an interferometric array is known as its $uv$-coverage, and defines the array's properties entirely.
For the VLA, for example, the instantaneous $uv$-coverage when observing the zenith will be as shown in Fig. \ref{fig.vla.uvcoverage}.

@@

\pg
Assuming that all the antennas in an array are equivalent and perfectly calibrated, the van Cittert-Zernike theorem (\citetads{1934Phy.....1..201V}, covered in \citetads{2001isra.book.....T}) allows us to equate a visibility with a single Fourier mode of the plane tangent to the sky where the instrument is pointed. This direction is called the \emph{phase centre}, because phase shifts are introduced between antenna voltages before averaging so as to point each visibility in this direction.

\pg
The Point-Spread Function (PSF) of an array is thus the inverse Fourier transform of the array's $uv$-coverage. The more element in the array, the better the $uv$-coverage and thus the better the PSF. An interferometer's PSF will always be worse than an antenna dish of equivalent diameter, the UV-coverage of which is the inverse Fourier transform of a disc. This is because an interferometer behaves like a \emph{sparse} dish: the sparsity of our coverage manifests as very large point-spread function. The more baselines are available, the less sparse the UV-coverage, and so the better the PSF. This is shown in Fig. \ref{fig.vla.PSFvsUVcov}.
\begin{figure}[h!]
\centering
\resizebox{\hsize}{!}{\includegraphics{images/PSFvsUVcov.png}}
\caption{\label{fig.vla.PSFvsUVcov}Plot of the PSF associated to more and more complete UV-coverage (here, consisting of more and more elements of the VLA). Image credit: Rick Perley, 3GC4 lecture.\textcolor{red}{crop out garbage on top}}
\end{figure}

\pg
This PSF can be improved by assuming that the sky is constant over periods of time (supersynthesis) or constant over fractions of the total bandwidth. This improves UV-coverage by a factor $N_\mathrm{t}\times N_\nu$, the number of measurements in time and frequency respectively. The effect of supersynthesis on the PSF is shown in Fig. \ref{fig.vla.PSFsupersynthesis}.
\begin{figure}[h!]
\centering
\resizebox{\hsize}{!}{\includegraphics{images/{PSF_supersynthesis}.png}}
\caption{\label{fig.vla.PSFsupersynthesis}Effect of supersynthesis on the UV-coverage of the full VLA. Image credit: Rick Perley, 3GC4 lecture.}
\end{figure}

@@

\pg
In practice, performing a full deconvolution on large images (easily over 25 million pixels in total) is not computationally viable. Scientists therefore resort to algorithms and techniques to accelerate imaging and deconvolution. For example, the use of Direct Fourier Transforms (DFT), which are extremely slow, is avoided when going from visibility space to image space - Fast Fourier Transforms (FFTs) are preferred. Similarly, one seeks to avoid having to perform full deconvolution on the dirty image, since many pixels in the image do not actually contain astrophysical signal. In this chapter, we will briefly cover the most common method of deconvolution used in radio astronomy.

@@

\pg
In Fig. \ref{fig.vla.clean}, we show the effect of deconvolution on a dirty image of 3C295 (a very bright \& complex source which is the subject of much of this thesis' work). Since this source includes diffuse emission, standard CLEAN is insufficient (since CLEAN assumes a sparse distribution of point-like sources). Instead, the DDFacet software package's novel MetroCLEAN algorithm (see \citetads{2017arXiv171202078T}), which deconvolves islands consisting of sets of pixels at once, was used. This allowed for a far greater dynamic range than standard CLEAN could achieve.

\pg
The problem of imaging radio interferometric data, of which deconvolution is but a subset, is still a very active field of research today. For more information on this topic, the NRAO website\footnote{See \href{https://www.cv.nrao.edu/~abridle/deconvol/node8.html\#SECTION00051000000000000000}{the CLEAN section of the NRAO website}.} is an excellent start for a general introduction to this topic, while \citet{2017arXiv171202078T} provides an outstanding entry point for cutting-edge work at the time of writing.

\section{Weighting Schemes}
\pg
In \cref{section.clean}, we discussed the problem of deconvolution in radio interferometric imaging. One consideration that is not raised is the issue of conditioning. Consider two extreme cases: a field containing a few point sources(\cref{fig.point}), and a field containing a diffuse and turbulent source, with very complex flux distribution at all scales (\cref{fig.cyga}). Deconvolving the PSF from the first image using CLEAN algorithms will be child's play, but doing the same with the second will be extremely complex. The first case is said to be well-conditioned, and the second to be poorly-conditioned.

@@

\pg
The better the conditioning, the easier the deconvolution and the closer to the ground truth its result. As such, when the true underlying structure of a source is poorly-known, getting the best conditioning possible for deconvolution becomes a key concern. Thankfully, because the actual measurements taken with interferometers are the Fourier modes of the sky brightness distribution, they can be weighted when reconstructing the images to fulfill specific requirements. In particular, Briggs weighting \citepads{1995AAS...18711202B} gives a sliding parameter between \emph{natural} weighting, where each visibility is weighed equivalently (this optimises noise in the field) and \emph{uniform} weighting, which gives each visibility a weight inversely proportional to $uv$-plane density. Uniform weighting optimises for resolution at the cost of image signal-to-noise.

\begin{figure}[h!]
\centering
\begin{subfigure}{.48\textwidth}\resizebox{\hsize}{!}{\includegraphics{images/natural.png}}
\caption{\label{fig.natural} Image made with LOFAR data and natural weighting.}
\end{subfigure}
\hfill
\begin{subfigure}{.48\textwidth}
\resizebox{\hsize}{!}{\includegraphics{images/uniform.png}}
\caption{\label{fig.uniform} Image made with the same data as \cref{fig.natural} and uniform weighting.}
\end{subfigure}
\hfill
\begin{subfigure}{.48\textwidth}
\resizebox{\hsize}{!}{\includegraphics{images/briggs0.png}}
\caption{\label{fig.briggs0} Image made with the same data as \cref{fig.natural} and Briggs weighting, with a Briggs parameter of 0.}
\end{subfigure}
\caption{\label{fig.briggs-weighting}Impact of different weighting schemes on interferometric image reconstruction. All the images shown above are deconvolved.}
\end{figure}

@@

\pg
Note that the weighting schemes shown here rely only on a choice between normalising visibilities by local $uv$-density vs. signal-to-noise, with Briggs weighting providing a formal method for going from one regime to another or optimising between both. Other weighting schemes exist - one can attempt to optimise data size vs decorrelation\footnote{As data is averaged, decorrelation becomes more and more important in the field. See \cref{section.RIME} for more details.}, for example \citepads[see]{2016MNRAS.462.2542A}. This can help ensure a homogeneous conditioning of the deconvolution problem in the image, by ensuring that the PSF is as peaked as possible throughout the field of interest. Part of the work of this thesis was in devising a weighting scheme dependent on calibration quality, to drastically improve deconvolution conditioning and final image quality. This weighting scheme is described in much greater detail in \cref{chapter.paper}.

\pg
We close this section by noting that data flagging \citepads[cf.]{2010ascl.soft10017O} can be considered a form of weighting scheme. This consists of assigning a null weight to those data points considered too corrupted by various instrumental or atmospheric effects (typically, Radio Frequency Interference - RFI) to be scientifically useful. Those weights which are not deemed too corrupted are assigned a unit weight. This corresponds to simply dropping those data points which would corrupt the image reconstruction by introducing unphysically strong fringes in the image.

### 2018-06-01 source/Ch2-intro/intro.tex
reorganised source files
added a lot

\chapter{Foreword}
\pg
write introduction to the thesis

@@

\pg
We start by linking interferometry to more concrete concepts: specifically, we will give a (very!) brief introduction to radio antennas and their characteristics. We will then use this introduction to contextualise the advantages and drawbacks of radio interferometry, along with the concrete technical problems that the method introduces.

@@

\clearpage
\input{source/intro/imaging.tex}
\clearpage
\input{source/intro/calibration.tex}
\clearpage
\input{source/intro/science.tex}
\clearpage
\input{source/intro/lofar.tex}

>>>>

@@

\pg
We start by linking interferometry to more concrete concepts: specifically, we will give a (very!) brief introduction to radio antennas and their characteristics. We will then use this introduction to show the advantages and drawbacks of radio interferometry, along with the concrete technical problems that the method introduces.

@@

\clearpage
\input{source/Ch2-intro/imaging.tex}
\clearpage
\input{source/Ch2-intro/calibration.tex}
\clearpage
\input{source/Ch2-intro/science.tex}
\clearpage
\input{source/Ch2-intro/lofar.tex}
